---
title: "CRAN History EDA"
author: "R Sangole"
output: html_notebook
---

```{r libraries, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(lattice)
library(naniar)
```

# Introduction



# Data Preparation {.tabset}

## Read In {.tabset}

```{r}
input_dir <- fs::path("../input")
files <- fs::dir_ls(input_dir)
his_dt <- read_csv(
  files[1],
  col_types = cols(
    package = col_character(),
    version = col_character(),
    date = col_date(format = "%Y-%m-%d"),
    repository = col_character()
  )
)
ov_dt <- read_csv(
  files[2],
  col_types = cols(
    package = col_character(),
    version = col_character(),
    depends = col_character(),
    imports = col_character(),
    license = col_character(),
    needs_compilation = col_logical(),
    author = col_character(),
    bug_reports = col_character(),
    url = col_character(),
    date_published = col_date(format = "yyyy-mm-dd"),
    description = col_character(),
    title = col_character()
  )
)
```

## Quick View {.tabset}

```{r}
dplyr::glimpse(his_dt, 100)
```

```{r}
dplyr::glimpse(ov_dt, 100)
```


## Conclusion




# Data Quality

```{r}
naniar::vis_miss(his_dt)
```

```{r}
ov_dt |> 
  dplyr::arrange(date_published) |> 
  vis_miss()
```

# Questions

* How long did packages take from their first release to version 1.0? 
* What type of packages were most frequent in different years? 
* Who are the most productive authors? 
* Can you predict the growth toward 2025?

# Temporal Exploration


* How long did packages take from their first release to version 1.0? 

```{r}
ov_dt |> 
  select(package, version)
```




# Sets

